---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
- Load data by read.csv and store it into `df_raw`
```{r, echo=TRUE}
df_raw <- read.csv('./activity.csv', header=TRUE)
```

- Remove `NA` values and store it into `df`
```{r, echo=TRUE}
df <- df_raw[!is.na(df_raw$steps),]
```

## What is mean total number of steps taken per day?
- Aggregate steps data by date and store into `df_by_day`
```{r, echo=TRUE}
df_by_day <- aggregate(steps~date,data=df,sum,na.rm=TRUE)
```

- Draw histogram 
```{r, echo=TRUE}
hist(df_by_day$steps, main="Histogram of mean total number of steps by day",xlab="# of Steps")
```

- Calculate the mean and median of the total number of steps taken per day
```{r, echo=TRUE}
mean_steps <- mean(df_by_day$steps)
median_steps <- median(df_by_day$steps)

```

Mean of total number of steps taken per day is **`r format(mean_steps,nsmall=2)`**.
Median of total number of steps taken per day is **`r format(median_steps,nsmall=2)`**.


## What is the average daily activity pattern?

- Caculate and plot the average daily activity pattern. First aggregate average steps by interval into `df_interval`, then plot the relationship of steps and interval. 
```{r, echo=TRUE}
df_interval <- aggregate(steps~interval,data=df,mean,na.rm=TRUE)
plot(steps~interval,data=df_interval,type="l")
max_interval <- df_interval[which.max(df_interval$steps),]$interval
```


On average across all the days in the dataset, the interval that contains the maximum number of steps is **``r max_interval``**.

## Imputing missing values

- Calculate and report the total number of missing values in the dataset
```{r, echo=TRUE}
num_of_NAs <- sum(is.na(df_raw$steps))
```

The total number of missing values is **``r num_of_NAs``**.

- Create a new dataset `df_fill` and fill in missing `NA`s.  The average steps of the corresponding interval is used to replace the missing value.

```{r, echo=TRUE}
df_fill <- df_raw
for (i in 1:nrow(df_fill)){
  if (is.na(df_fill[i,]$steps)){
    df_fill[i,]$steps = df_interval[df_interval$interval==df_fill[i,]$interval,]$steps
  }
}
```

- Aggregate steps data by date and store in `df_fill`
```{r, echo=TRUE}
df_fill_by_day <- aggregate(steps~date,data=df_fill,sum,na.rm=TRUE)
```

- Draw histogram 
```{r, echo=TRUE}
hist(df_fill_by_day$steps, main="Histogram of mean total number of steps by day",xlab="# of Steps")
```

- Calculate the mean and median of the total number of steps taken per day
```{r, echo=TRUE}
mean_steps_fill <- mean(df_fill_by_day$steps)
median_steps_fill <- median(df_fill_by_day$steps)
```

Mean of total number of steps taken per day is **`r format(mean_steps_fill,nsmall=2)`**.
Median of total number of steps taken per day is **`r format(median_steps_fill,nsmall=2)`**.

After imputing missing values, the median value of total number of steps taken per day increases slightly and the mean value does not show noticeable difference.

## Are there differences in activity patterns between weekdays and weekends?

```{r, echo=TRUE}
df_fill$weekday <- ifelse(weekdays(as.Date(df_fill$date))=="Saturday"|weekdays(as.Date(df_fill$date))=="Sunday","weekend","weekday")
df_fill$weekday <- factor(df_fill$weekday,levels=c("weekday","weekend"))
```

```{r, echo=TRUE}
library(lattice)
df_interval_2 <- aggregate(steps~interval+weekday, df_fill, mean)
xyplot(steps~interval|factor(weekday), data=df_interval_2, aspect=1/2, type="l")
```

The plot indicates that 

- In weekdays, a peak of steps exists around interval 700-800.

- In weekend, activities between interval 1100-1700 is higher than weekdays.